Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-04 papers/actus2021/article.tex
comments
more fixes
Revised to end section 7.

\textbf{DONE:}
\begin{enumerate}
\item
Added overview at the end of the intro.
\item
Tidying up identifiers in math equations using the command \verb+\ident{...}+
\item
Biblio sorted out ordering (added keys). shortened author and editor lists
\item
Revised sections 2--4.
\end{enumerate}

@@

% SJT: added this overview of the paper
% it is traditional to end the introduction like this,
% 'signposting' what is in the rest of the paper.
Section \ref{background} of the paper covers the relevant financial background, and Section \ref{modelling} introduces Marlowe as well as the ACTUS financial standard. Section \ref{executable} builds an executable specification of ACTUS in Haskell, and this is used in Section \ref{generation} to generate  the Marlowe code for an ACTUS contract from the terms of the contract. Section \ref{tokenization} explains how tokens are used to represent ownership of roles in a running contract, and Section \ref{verification} presents how we provide assurance of the contract correctness. Section \ref{related} examines related work and Section \ref{conclusion} concludes.

@@

Every participant of an ACTUS contract is described by a Role which
is in its turn represented through a special unique non-fungible token.
This makes every Marlowe contract a tradable security, allowing a
participant to sell its ``share'' in a contract by selling a corresponding
role token.

Role tokens can potentially allow more complex manipulation over shares,
especially when share represents an incoming cash flow: participants
only send funds to a party represented by a given token. Such a token
would represent a positive cashflow which in turn could not just become
tradable but would also allow to derive tokens only representing fractions
of a particular cash flow in a contract.

Moreover, it turns Actus-loans into derivatives, for example contracts
like Interest Rate Swap (and Swaps in general) could be approximated
by an Atomic Swap of tokens representing incoming cash flows from
lending.

For example, if Alice has fixed income from a loan (or some other
investment) and Bob has comparable but variable (fluctuating) income
- Bob can hedge by swapping cash flows with Alice. If Alice's income
is locked with token1 and Bob's income is locked with token2 - then
atomic swap of those tokens is equivalent to a swap of cash flows.

\section{Verification and testing}
\label{verification}

@@

In order to test the equivalence between Executable Haskell implementation
of ACTUS for smart-contracts and original Java implementation for
analytics - a simple property test was introduced:

\smallskip
\noindent 
\begin{equation}
\forall ct\forall rf.getCashFlows("haskell",ct,rf)\equiv getCashFlows("java",ct,rf)
\end{equation}

Where $ct$ represents contract terms, $rf$ is risk factor model and
$getCashFlows()$ returns a set of $(date,payoff)$ tuples.

@@

QuickCheck\cite{qc} contract terms generators also allow to ensure
arbitrary properties of a Marlowe contract. This could be enhanced
with Marlowe's static analysis feature by utilising Assert operator:

\begin{verbatim}
ContractTerms := sample(qcgenerator) 
Contract := generateMarloweContract(ContractTerms) 
ContractWithAssert := appendAssertion(Contract, assertion) 
runStaticAnalysis(ContractWithAssert)
\end{verbatim}

While it doesn't cover all possible contracts it could guarantee that
property holds for a statistically significant fraction of a contract.
Using Static Analysis to check a particular contract instead of randomly
sampling allows: 
\begin{itemize}
\item more balanced sampling: Space of contract terms depends linearly on
the coverage, e.g. if we cover 10\% of all possible contract terms
- we'll cover 10\% of all contracts. Different contracts have different
amount of risk factors thus different search space, for instance a
contract with 3 risk factors (e.g. 3 observations of an interest rate)
would span over n\textasciicircum 3 while a contract with 10 risk
factors would span over n\textasciicircum 10. Meanwhile, space of
all risk factors in all contracts doesn't linearly depend on the coverage. 
\item dependency tracking: SMT-solver would likely be more aware of execution
paths leading to failure of a test, which could significantly reduce
search space 
\item completeness: If not timed out, SMT-solver is decidable while sampling
is semi-decidable. 
\end{itemize}

@@

By design, the Marlowe interpreter always refunds assets held in a
contract when it terminates. This is done in order to ensure that
no funds are lost forever. The funds are returned to whoever's internal
account actually holds them.

However, this presents as a problem in certain cases where ownership
of the funds could not be automatically determined. For example if
Alice puts collateral in a crypto-loan contract - she would formally
maintain ownership, which means she would get automatically refunded
when Close contract is reached.

This implicit refund is easy to overlook by smart-contract developers
as plain Close is often used as a default action in case of unexpected
behaviour like timeouts(especially Choice timeout). This can easily
lead to multi-million dollar mistakes if malicious Alice decides to
exploit an auto-refund feature in order to get her collateral without
paying back. Here's the example of simple attack vector: 
\begin{enumerate}
\item Alice creates a contract where Deposit timeout would lead to Close 
\item Bob doesn't know or test the timeout path. Even if Bob is a programmer
- he might not be aware that Closing contract would cause collateral
refund 
\item Alice puts collateral and gets notional from Bob 
\item Alice doesn't pay for a loan (Deposit timeout) 
\item Alice gets her collateral back 
\item Bob loses notional 
\end{enumerate}
In order to prevent this from happening, an additional ''Auto-Refund''
security check was introduced as part of Static Analysis tooling which
notifies users about all ``Close'' constructs that can lead to automatic
refunds and encourages users to write explicit logic for edge cases.

>>>>

\textbf{DONE:}
\begin{enumerate}
\item
Added overview at the end of the intro.
\item
Tidying up identifiers in math equations using the command \verb+\ident{...}+
\item
Biblio sorted out ordering (added keys). shortened author and editor lists
\item
Revised sections 2--7.
\end{enumerate}

@@

A note on notation: \texttt{typewriter} font will be used for Marlowe constructs
% SJT and fragments?
while \emph{math} font will be used for mathematical formulas and pseudo-code.
% Have I got this right?

% SJT: added this overview of the paper
% it is traditional to end the introduction like this,
% 'signposting' what is in the rest of the paper.
Section \ref{background} of the paper covers the relevant financial background, and Section \ref{modelling} introduces Marlowe as well as the ACTUS financial standard. Section \ref{executable} builds an executable specification of ACTUS in Haskell, and this is used in Section \ref{generation} to generate  the Marlowe code for an ACTUS contract from the terms of the contract. Section \ref{tokenization} explains how tokens are used to represent ownership of roles in a running contract, and Section \ref{assurance} describes how we provide assurance that  contracts behave as they should. Section \ref{related} examines related work and Section \ref{conclusion} concludes.

@@

Every participant of a Marlowe contact %ACTUS contract 
is described by a \texttt{Role} which
is in its turn represented through a unique non-fungible token, created at the time that the contract instantiated on the blockchain.
This makes every ACTUS contract a tradable security, allowing a
participant to sell its \emph{share} in a contract by selling a corresponding
role token.

Role tokens can potentially allow more complex manipulation over such shares,
especially when the share represents an incoming cash flow; in that case, participants
send funds to a party represented by a given token. Such a token
would represent a positive cashflow, which in turn could not only become
tradable but could also allow the derivation of tokens representing \emph{fractional parts}
of a particular cash flow in a contract.

Moreover, this process turns ACTUS loans into derivatives. For example, contracts
like \emph{Interest Rate Swap} (and \emph{Swaps} in general) could be approximated
by an \emph{Atomic Swap} of tokens representing incoming cash flows from
loans. 
For example, if Alice has fixed income from a loan, or some other
investment, and Bob has comparable but variable (fluctuating) income,
Bob can hedge by swapping cash flows with Alice. If Alice's income
is locked with \texttt{token1} and Bob's income is locked with \texttt{token2} then an
atomic swap of those tokens is equivalent to a swap of cash flows.

\section{Assurance}
\label{assurance}

We provide assurance to users of our Marlowe ACTUS contracts in a number of ways, as described here.

@@

First, we are able to test the executable Haskell implementation
of ACTUS for smart-contracts by comparing it with an existing implementation written in Java. To do this  
%original Java implementation for analytics - 
a simple property-based test in QuickCheck~\cite{qc} was introduced, where we generate contract terms and risk factors randomly to test the property.

\newcommand{\rf}{\ident{rf}}
\newcommand{\ct}{\ident{ct}}
\smallskip
\noindent 
\begin{equation}
\forall \ct.\forall \rf. \; \ident{getCashFlows}(\ident{"haskell"},\ct,\rf)\equiv \ident{getCashFlows}(\ident{"java"},\ct,\rf)
\end{equation}
\noindent
where \emph{ct} represents contract terms, \emph{rf} is risk factor model, and
\emph{getCashFlows()} returns a set of \emph{(date,payoff)} tuples.

@@

QuickCheck contract terms generators also allow us to check
other properties of a Marlowe contract. This could be enhanced
with Marlowe's static analysis feature by utilising the \texttt{Assert} operator:

% SJT what language is this written in?
% can we turn it into Haskell? or Java?
\begin{verbatim}
ContractTerms := sample(qcgenerator) 
Contract := generateMarloweContract(ContractTerms) 
ContractWithAssert := appendAssertion(Contract, assertion) 
runStaticAnalysis(ContractWithAssert)
\end{verbatim}

While this scenario does not cover all possible contracts, it could guarantee that
property holds for a statistically significant fraction of a contract.

\subsection{Static analysis for verification}
% SJT added this as a separate sub section, as it covers the static analysis work.
% references added.

Using static analysis for Marlowe~\cite{} it is possible fully to check a particular contract instead of using random
sampling. That would allow some refinements: 
\begin{itemize}
\item More balanced sampling: the space of contract terms depends linearly on
the coverage, e.g. if we cover 10\% of all possible contract terms
- we'll cover 10\% of all contracts. Different contracts have different
sets of risk factors, and so different search spaces. For instance, a
contract with 3 risk factors (e.g. 3 observations of an interest rate)
would span over $n^3$ values while a contract with 10 risk
factors would span over $n^{10}$. Meanwhile, the space of
all risk factors in all contracts doesn't linearly depend on the coverage.
% SJT explain last sentence. 
\item Dependency tracking: an SMT-solver is potentially likely to be more aware of execution
paths that lead to the failure of a test, a feature that could significantly reduce
search space. 
\item Completeness: if not timed out, SMT-solving is decidable while sampling
is semi-decidable. 
\end{itemize}

@@

By design, the Marlowe interpreter always refunds any assets held in a
contract when it terminates. This is done in order to ensure that
no funds are lost forever. The funds are returned to whoever's internal
account  holds them.

However, this presents as a problem in certain cases where ownership
of the funds could not be  determined automatically. For example, if
Alice puts collateral in a crypto-loan contract she would formally
maintain ownership, which means she would get automatically refunded
when a \texttt{Close} construct is reached.

This implicit refund is easy to overlook by smart-contract developers
as plain \texttt{Close} is often used as a default action in case of unexpected
behaviour like timeouts, and especially a \texttt{Choice} timeout. This can easily
lead to costly %multi-million dollar 
mistakes if  Alice maliciously decides to
exploit an auto-refund feature in order to get her collateral without
paying back, as in the following scenario:
\begin{enumerate}
\item Alice creates a contract where a \texttt{Deposit} timeout would lead to \texttt{Close} 
\item Bob doesn't know or test the timeout path. Even if Bob is a programmer,
he might not be aware that \texttt{Close}ing the contract would cause the collateral
to be refunded to Alice.
\item Alice puts her collateral in the contract and gets the notional from Bob.
\item Alice doesn't pay for the loan: i.e. there is a \texttt{Deposit} timeout. 
\item Alice gets her collateral back. 
\item Bob loses his notional. 
\end{enumerate}
In order to prevent this from happening, an additional \texttt{Auto-Refund}
security check was introduced as part of static analysis tooling which
notifies users about all \texttt{Close} constructs that can lead to automatic
refunds and encourages users to write explicit logic for edge cases.
